# Final Project
GEOG693 Final Project

Sara Schreder-Gomes

-------------------------

#### Install packages
The user must install the following packages: 

tidyverse, base, ggplot2, plotly, knitr, kableExtra, sf, rnaturalearth, maps, grid
```{r, libraries, include =FALSE}
library("tidyverse")
library("base")
library("ggplot2")
library("plotly")
library("knitr")
library("kableExtra")
library("sf")
library("rnaturalearth")
library("maps")

library("oz") #install necessary packages for site map
library("grid")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Enter Field data

I had to do some manual data entry in Excel to set up the field data in the format I wanted it to be in. I decided to do this instead of trying to import from a PDF and manually manipulating from there.

```{r}
FieldData <- data.frame("Sample" = c(NA, NA, NA, NA, "KCB 1", "KCB 2", "KCB 3", "KCB 4", "KCB 5", "KCB 6", "KCB 7", "KCB 8", "KCB 9", "KCB 10", "KCB 11"),
           "Location" = c(NA, NA, NA, NA, "Gneiss Lake", "Lake Magic I", "Lake Magic II", "Lake Aerodrome", "Twin Lake West", "Salar Gorbea site 7", "Salar Gorbea site 10","Salar Ignorado site 3", "Saline Valley", "Bonneville Salt Flats", "Crosbie Lake"),
           "Country" = c(NA, NA, NA, NA, "Australia", "Australia", "Australia", "Australia", "Australia", "Chile", "Chile", "Chile", "USA", "USA", "Canada"),
           "Year Collected" = c(NA, NA, NA, NA,2015, 2015, 2015, 2009, 2009, 2009, 2007, 2007, NA, NA, NA),
           "Water Depth (cm)" = c(NA, NA, NA, NA,2.5, 2.5, 5, 1.2, 10, 1, 2, 10, NA, NA, NA),
           "Water Color" = c(NA, NA, NA, NA,"Yellow","Yellow", "Yellow", "Yellow", "Green", "Clear", "Pink", "Clear", NA, NA, NA),
           "pH" = c(NA, NA, NA, NA,1.4, 2, 1.6, 1.8, 2.6, 2.7, 1.8, 3.8, NA, NA, NA),
           "Salinity (%TDS)" = c(NA, NA, NA, NA,32, 30, 30, 29, 27, 28, 18, 5, NA, NA, NA),
           "Latitude" = c(NA, NA, NA, NA,-33.21984, -32.43031 ,-32.42846, -32.20949, -33.05573, -25.42514, -25.42161, -25.49993,NA,NA,NA),
           "Longitude" = c(NA, NA, NA, NA,121.75405, 118.90505, 118.90552, 121.76072, 121.67550, -68.68759, -68.69366, -69.62456,NA,NA,NA),
            stringsAsFactors = FALSE)

write.csv(FieldData, "data/FieldData.csv", row.names = FALSE)

FieldData %>%
  kable() %>%
  kable_styling()
```


#### Munge the raw lab data

```{r}
LabData <- as.data.frame(read.csv("data/BrineCompostions_ActLabsResults_RAWData.csv", stringsAsFactors = FALSE), header= FALSE, sep= ",", dec = ".", col.names = NULL) 

# for any value containing "<" (indicating uncertainty in the measurement), replace value with NA
for (col in colnames(LabData)){
    is.na(LabData[[col]]) <- startsWith(LabData[[col]], "<")
}

# exclude columns with no analyte data
for (col in colnames(LabData)){
    if (col == "Sample" || col == "Analyte.Symbol" || col == "Date.Time") 
        next
    
# define multiplier to convert units
    multiplier <- 1
    if (LabData[1,col] == "mg/L")
        multiplier <- .001
    
    # Define detection limit for an analyte
    limit <- as.numeric(LabData[2,col]) * multiplier
    
    #remove values that are less than multiplied value (detection limit)
    for(i in 1:nrow(LabData))
    {
        if (i < 3) 
            next
        if (is.na(LabData[i,col]) == FALSE && LabData[i,col] < limit)
            is.na(LabData[i,col]) <- TRUE
    }
}
```

#### Create data frames with selected elements for further analysis
```{r}
MajorIonLabData <- LabData %>% select(Sample,Analyte.Symbol, Na, Si, Fe, Al, Cl, Mg, Ca, K, Br, SO4) 
#choose the ion concentrations for analysis. This makes it easy to add analytes to the shorter data frame at a later point.

MajorIonLabData %>%
  kable() %>%
  kable_styling()
```

```{r}
REELabData <- LabData %>% select(Sample,Analyte.Symbol, Ce, Dy, Er, Eu, Gd, Ho, La, Lu, Nd, Pr, Sm, Sc,Tb, Tm, Yb, Y) 
#choose the ion concentrations for analysis. This makes it easy to add analytes to the shorter data frame at a later point.

REELabData %>%
  kable() %>%
  kable_styling()
```

```{r} 
#merge LabData and FieldData to one df 
MergedDatasets <- merge(FieldData, LabData, by.x = "Sample", by.y = "Sample")

MergedDatasets %>%
  kable() %>%
  kable_styling()
```
This merge does not include three samples from the lab data because they are not acid. They aren't necessary for further analysis with these research questions. 

#### Create scatter plots for element input

```{r}
# Create basic scatter plot

#   Make this a function
    #Usage: pHvs(elementAbbrev, concentration unit?)
    # throw error if abbrev isnt in the data. Message: 'Element abbreviation not found in data'

#Tried to make this into a function, but there's a small problem somewhere. 
#     pHvs <- function(Element){
#     TestPlot <- ggplot(MergedDatasets, aes(x=as.numeric(pH), y=as.numeric(Element))) +
#        xlab("pH") + ylab("Fe (mg/L)") + geom_point(size = 3, shape = 20) +
#        geom_text(aes(label=Sample), vjust=-.5) + ggtitle("pH vs. Fe content for Acid Saline Lake Systems")
#return(TestPlot)
#}

# I am under the gun grading, but would be happy to work this out with you next semester.  Would need to update the title to reflect element choice as well.  But should be possible.

ggplot(MergedDatasets, aes(x=as.numeric(pH), y=as.numeric(Fe))) + xlab("pH") + ylab("Fe (mg/L)") + geom_point(size = 3, shape = 20) + geom_text(aes(label=Sample), vjust=-.5) + ggtitle("pH vs. Fe content for Acid Saline Lake Systems")
#by setting aes(x and y) as.numeric, it fixes a data coersion problem. In previous coercion, they were set to 'chr', which led to samples with 'NA' values being plotted and a very "wonky" y-axis.

ggplot(MergedDatasets, aes(x=as.numeric(pH), y=as.numeric(Si))) + xlab("pH") + ylab("Si (mg/L)") + geom_point(size = 3, shape = 20) + geom_text(aes(label=Sample), vjust=-.5) + ggtitle("pH vs. Si content for Acid Saline Lake Systems")

ggplot(MergedDatasets, aes(x=as.numeric(pH), y=as.numeric(Al))) + xlab("pH") + ylab("Al (mg/L)") + geom_point(size = 3, shape = 20) + geom_text(aes(label=Sample), vjust=-.5) + ggtitle("pH vs. Al content for Acid Saline Lake Systems")
```

#### Create line plot of rare earth elements

Goal: see if there are any correlations between pH and the occurance of rare earth elements

ex: a straight (ish) line might indicate that the occurance of that element might be controlled by pH helps answer the research question: How does pH affect water chemistry?

```{r}
# sort pH in ascending order
MergedDatasets <- MergedDatasets[order(MergedDatasets$pH),] 

#plot selcted rare earth elements vs. pH as line plot. More traces can be added. 
pHvsRareEarthElements <- plot_ly(MergedDatasets, x = as.numeric(MergedDatasets$pH), y = as.numeric(MergedDatasets$Gd), name = 'Gd', type = 'scatter', mode = 'lines',
        line = list(color = 'rgb(205, 12, 24)', width = 2)) %>%
  add_trace(y = MergedDatasets$Dy, name = 'Dy', line = list(color = 'rgb(22, 96, 167)', width = 2)) %>% #add another element
  add_trace(y = MergedDatasets$Yb, name = 'Yb', line = list(color = 'rgb(205, 12, 24)', width = 2, dash = 'dash')) %>%
  layout(title = "Selected Rare Earth Elements vs. pH",
         xaxis = list(title = "pH"),
         yaxis = list (title = "Concentration (ug/L)"))
pHvsRareEarthElements
```

#### Study site map

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)

#create dataframe with site latitude and longitude
site_id <- c("Gneiss Lake", "Lake Magic I", "Lake Magic II", "Lake Aerodrome", "Twin Lake West")
    siteLongs <- c(121.75405, 118.90505, 118.90552, 121.76072, 121.67550)
    siteLats <- c(-33.21984, -32.43031 ,-32.42846, -32.20949, -33.05573)
    sites <- data.frame(site_id = site_id, longitude = siteLongs, latitude = siteLats)

#set coordinate system and sites
sites <- st_as_sf(sites, coords = c("longitude", "latitude"), 
    crs = 4326, agr = "constant")

#create study site map
SiteMap <- ggplot(data = world) +
    geom_sf() +
    geom_sf(data = sites, size = 4, shape = 16, color = "darkred") +
    coord_sf(xlim = c(113, 127), ylim = c(-29, -36), expand = TRUE) +
    geom_sf_label(data=sites, aes(label = site_id), nudge_x = 1.2, nudge_y = .4) +
    coord_sf(xlim = c(113, 127), ylim =c(-29, -36), expand = TRUE)

#Create inset map
AustraliaMap <- ggplot(data = world) + geom_sf() + geom_sf(data = sites, size = 4, shape = 16, color = "darkred") + coord_sf(xlim = c(111, 156), ylim = c(-41, -7)) 

AustraliaMap
```

```{r}
#Create combined map with inset
library(grid)
grid.newpage()
vp_b <- viewport(width = 1, height = 1, x = 0.5, y = 0.5)  # the larger map
vp_a <- viewport(width = 0.3, height = 0.3, x = 0.85, y = 0.25)  # dimentions and location of inset in the lower right 
print(SiteMap, vp = vp_b)
print(AustraliaMap, vp = vp_a)
```
#well done on the inset!
